Experience-Biased Symbolic Planning

### 2014-05-13 AAAI_AuthorKit/paper.tex
Mike's Edits Part 2, now 4 pages done

We may view planning as a process that generates distance-to-goal labels for each node of a graph.
Given these distance labels, it's a simple matter to generate a plan by tracing steps from the start, successively taking the optimal successor.
Since there is no worst-case polynomial-time algorithm that can solve the shortest-paths problem in STRIPS (recall the graph is exponential-sized), we may try to roughly approximate the distance labels.

The approximate labels we call \textbf{heuristics}, and we use them in hopes of finding plans more efficiently.
If the heuristic precisely reflects the true distances, we can immediately retrieve a path.
Algorithms such as A* are designed to find solutions quickly whenever the heuristic is of reasonable quality.
If we hope to take subexponential time, we can only afford to examine a vanishing fraction of the states, and it is only for these that we compute the heuristic.
If the heuristic is good, we hope not to have to look at (i.e. generate) too many states.
Hence, there is a tradeoff between spending more time to compute a more accurage heuristic, vs evaluating a weaker heuristic quickly but at many more states.

In practice, even using state-of-the-art methods to generate heuristics on isolated planning instances, STRIPS problems can be very difficult to solve without domain knowledge.
Even humans have difficulty when faced with an unfamiliar kind of puzzle, though we get better with experience.
Thus, one might hope that a planning agent would similarly learn to generalize solutions from past planning experiences to related new instances.
A recent approach builds \textbf{Experience graphs} to estimate the high-level connectivity of the free space in motion planning tasks \cite{phillips2012graphs}.
The intuition is to remember previously generated paths so that, when a new start and goal is queried, a new path can be quickly generated by reusing subpaths from the E-graph.
However, E-graphs have never before been applied to high-level representations such as STRIPS. This is our present contribution.

In section [???], we present the HSP heuristic for STRIPS.
Then, we discuss E-graphs in more technical detail.
This paper's novel contributions follow thereafter, where we extend E-graphs to STRIPS planning by using it to augment the HSP heuristic. 
The resulting A* search is complete with a configurable suboptimality guarantee.
Finally, we present experiment results in which this use of E-graphs is demonstrated to achieve considerable speedups over experience-free HSP in several planning domains.

@@

The problem $P$ defines a directed state graph $(V,c)$ where $V$ is the power set of $A$ (i.e. states $S\in V$ correspond to collections of atoms), and the edge costs are
\begin{eqnarray*} c(S,S') = \min\{Cost(op) \mid S\subseteq Prec(op)\text{ and}
\\S' = \left(S \setminus Del(op)\right) \cup Add(op)\} \end{eqnarray*}
%consists of states transitions $(u,v)$ for which there exists an operator $op\in O$ with $u\in Prec(op)$ and $v = u - Del(op) \cup Add(op)$; the weight of this edge is the minimum cost of such an $op$.
By default, $c(S,S') = \infty$ when there is no operator directly transitioning from $S$ to $S'$.
Given a STRIPS problem $P$, we seek a low-cost path from the initial state $I\in V$ to any of the goal states $S\in V$ such that $S \supset G$.

@@

\begin{defn} For $\epsilon\ge 1$, a heuristic $h$ is called \textbf{$\epsilon$-consistent} if $h(S) = 0$ when $S$ is a goal and $h(S) \le \epsilon c(S,S') + h(S')$ for all $S,S'\in V$. $h$ is \textbf{consistent} if it is $1$-consistent. \end{defn}
\begin{thm} Given an $\epsilon$-consistent heuristic and expanding no node more than once, A* is guaranteed to find a path costing no more than $\epsilon$ times the optimal path cost.  \end{thm}

@@

In the limit as $\epsilon^E \rightarrow 1$, following E-graph edges offers no advantage as $h$ is already an underestimate. Hence, using the triangle inequality to coalesce the sum,
\[h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N h(S_{i-1},S_i) = \min_{S_N} h(S_0,S_N)\]

@@

We cannot afford to compute $h^E$ according to its literal definition, as there are far too many paths to consider. Fortunately, there exists a practical means of computing it. By the triangle inequality, any consecutive pair of $h(S_{i-1},S_i)$ terms can be merged into one. Thus we lose no generality in restricting $S_i$ to lie on the E-graph for $0 < i < N$.

Let $V^E$ be the union of the goal states and the E-graph's vertices.
In addition to following E-graph edges, we imagine it's possible to ``jump" from $S$ to $S'$ with cost $\epsilon^E h(S,S')$. That is, let $c'(S,S') = \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$

In a preprocessing stage before the main search, we apply Dijkstra's algorithm once in reverse with the costs $c'$ to compute the estimated distance-to-goal $h^E(S)$ from every $S\in V^E$. If the heuristic $h$ is cheap to compute, this preprocessing takes $O(|V^E|^2)$ time, which is insignificant for small E-graphs and goal sets.

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which takes O($|V^E||A|$) time. A pseudocode implementation is listed in Algorithm \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while $\epsilon$ provides an additional goal-directed bias.

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\STATE Let $c(S,S')$ for $S,S'\in V^E$ represent E-graph edge costs.
\FORALL{$S \in V^E$}
\STATE ComputeG($S$)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated by A*}
\STATE ComputeG($S$)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG($S$) implements the dynamic programming computation of $g_S$-values.
Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations.
Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound on Algorithm \ref{alg:ComputeG} is $O(\mathbf{A^2O})$.
Thus, each state generated costs $O(\mathbf{A^2O + V^EA})$ time.
The preprocessing cost is $O(\mathbf{V^EA^2O + (V^E)^2A})$, asymptotically equivalent to preemptively generating every state in $V^E$.

@@

Finally, we remark that the Experience-Biased HSP algorithm is highly parallelizable, even if A* is retricted to expanding nodes sequentially. Recall that the sum, min and max functions are computable in logarithmic time using linearly many processors. With minor adjustments, then, ComputeG($S$) has parallel depth $O(\mathbf{A\log A\log O})$. With $\mathbf{AO}$ processors, the total cost of generating a state is $O(\mathbf{(A\log O + \log V^E)\log A})$. Using even more processors, the cost of \textit{expanding} a state can be reduced to the same by generating its successors in parallel.

@@

\section{Analysis of HSPr (section might be deleted, but please have a look)}

@@

>>>>

In forward weighted A* search, a frontier of nodes expands from the start until it hits a goal, at which point a solution is found.
Rather than blindly searching the entire exponential-sized space, A* tries to focus toward the goal by expanding nodes which seem likely to lie on a short solution path.
The minimum length of a solution passing through a given frontier node is the sum of distances from the start to this node, and from this node to a goal.
The better we estimate these, the sooner we'll reach a solution.

Since A* expands in a fairly conservative fashion from the start state onward, the start-to-frontier estimates are computed quite accurately by construction of a valid path.
Thus, the principal issue is the derivation of frontier-to-goal estimates.
Given exact distances, it would be a simple matter to trace a valid path from the start, successively taking optimal successors until a goal is reached.
In practice, we are forced to use estimates, which sometimes mislead us into local optima, wasting time on dead-end nodes.

These distance approximations are called \textbf{heuristics}, and we use them in hopes of finding good plans more efficiently.
Algorithms such as A* are designed to find solutions quickly whenever the heuristic is of reasonable quality.
If we hope to take subexponential time, we can only afford to examine a vanishing fraction of the states, and it is only for these that we compute the heuristic.
If the heuristic is good, we hope not to have to visit (i.e. generate) too many states.
Hence, there is a tradeoff between spending more time to compute a more accurate heuristic, vs evaluating a weaker heuristic quickly but at many more states.

In practice, even using state-of-the-art methods to generate heuristics on isolated planning instances, STRIPS problems can be very difficult to solve without domain knowledge.
Even humans have difficulty when faced with an unfamiliar kind of puzzle, though we get better with experience.
Thus, one might hope that a planning agent would similarly learn to generalize solutions from past planning experiences to related new instances.
A recent approach builds \textbf{experience graphs} to estimate the high-level connectivity of the free space in motion planning tasks \cite{phillips2012graphs}.
The intuition is to remember previously generated paths so that, when a new start and goal is queried, a new path can be quickly generated by reusing subpaths from the E-graph.
However, E-graphs have never before been applied to high-level representations such as STRIPS. This is our present contribution.

In section [???], we present the HSP heuristic for STRIPS.
Then, we discuss E-graphs in more technical detail.
This paper's novel contributions follow thereafter, where we extend E-graphs to STRIPS planning by using it to augment the HSP heuristic. 
The resulting weighted A* search is complete and has a configurable suboptimality guarantee.
Finally, we present experiment results in which this use of E-graphs is demonstrated to achieve considerable speedups over experience-free HSP in several planning domains.

@@

The problem $P$ defines a directed state graph $(V,c)$ where $V$ is the power set of $A$ (i.e. states $S\in V$ correspond to collections of atoms), and the edge costs are
\begin{eqnarray*} c(S,S') = \min\{Cost(op) \mid S\supseteq Prec(op)\text{ and}
\\S' = \left(S \setminus Del(op)\right) \cup Add(op)\} \end{eqnarray*}
%consists of states transitions $(u,v)$ for which there exists an operator $op\in O$ with $u\in Prec(op)$ and $v = u - Del(op) \cup Add(op)$; the weight of this edge is the minimum cost of such an $op$.
By default, $c(S,S') = \infty$ when there is no operator directly transitioning from $S$ to $S'$.
Given a STRIPS problem $P$, we seek a low-cost path from the initial state $I\in V$ to any of the goal states $S\in V$ such that $S \supseteq G$. In this paper, we are particularly concerned with sequences of problems in which $A$ and $O$ are fixed, but various pairs $\langle I,G\rangle$ are queried.

@@

\begin{defn} For $\epsilon\ge 1$, a heuristic $h$ is called \textbf{$\epsilon$-consistent} if $h(S) = 0$ when $S$ is a goal and $h(S) \le \epsilon c(S,S') + h(S')$ for all $S,S'\in V$. $h$ is \textbf{consistent} if it is $1$-consistent. \end{defn}
\begin{thm} Given an $\epsilon$-consistent heuristic and expanding no node more than once, A* is guaranteed to find a path costing no more than $\epsilon$ times the optimal path cost. [CITE ARA*] \end{thm}

@@

In the limit as $\epsilon^E \rightarrow 1$, E-graph edges offer no advantage as $h(S_{i-1},S_i) \le c^E(S_{i-1},S_i)$. Hence, using the triangle inequality to coalesce the sum,
\[h^E(S_0) = \min_{N,S_N,\pi} \sum_{i=1}^N h(S_{i-1},S_i) = \min_{S_N} h(S_0,S_N)\]

@@

We cannot afford to compute $h^E$ according to its literal definition, as there are far too many paths to consider.
Fortunately, there exists a practical means of computing it. By the triangle inequality, any consecutive pair of $h(S_{i-1},S_i)$ terms can be merged into one.
Thus, we lose no generality in restricting $S_i$ to lie on the E-graph for $0 < i < N$.
The heuristic computation can focus on the E-graph, using $h(S_{i-1},S_i)$ to estimate the cost of leaving the E-graph at $S_{i-1}$ before re-entering it at $S_i$.

Let $V^E$ be the union of the goal states and the E-graph's vertices.
In addition to following E-graph edges, we imagine it's possible to ``jump" from $S$ to $S'$ at cost $\epsilon^E h(S,S')$.
That is, let $c'(S,S') = \min\left(c^E(S,S'),~\epsilon^E h(S,S')\right)$

In a preprocessing stage before the main search, we apply Dijkstra's algorithm once in reverse with the costs $c'$ to compute the estimated distance-to-goal $h^E(S)$ from every $S\in V^E$. Ignoring the computation of $h$ (presently considered as a blackbox), this preprocessing takes $O(|V^E|^2)$ time, which is insignificant for small E-graphs and goal sets.

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which takes O($|V^E||A|$) time. A pseudocode implementation is listed in Algorithm \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while the weight $\epsilon$ provides a uniform goal-directed bias.

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\FORALL{$S \in V^E$}
\STATE ComputeG($S$)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c^E(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated by A*}
\STATE ComputeG($S$)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG($S$) implements the dynamic programming computation of $g_S$-values.
Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations.
Multiplying nested loop iterations together and ignoring cardinality signs for brevity, the runtime bound on Algorithm \ref{alg:ComputeG} is $O(\mathbf{A^2O})$.
Thus, each state generated costs $O(\mathbf{A^2O + V^EA})$ time.

The preprocessing cost is $O(\mathbf{V^EA^2O + (V^E)^2A})$, asymptotically equivalent to preemptively generating every state in $V^E$.
In fact, most of the preprocessing requires no knowledge of $G$, so it can be done in advance of the new planning query. Once the query $\langle I,G\rangle$ is given, the remaining work consists of completing the $c'(S,G)$ computations from $g_S(a),a\in G$, and then running Dijkstra's algorithm from $G$. This costs only $O(\mathbf{V^E(A + V^E)})$ time.

@@

Finally, we remark that the Experience-Biased HSP algorithm is highly parallelizable, even if A* is retricted to expanding nodes sequentially. Recall that the sum, min and max functions are computable in logarithmic time using linearly many processors. Therefore, with minor adjustments, ComputeG($S$) has parallel depth $O(\mathbf{A\log A\log O})$. With $\mathbf{A\max(O,V^E)}$ processors, the total cost of generating a state is $O(\mathbf{(A\log O + \log V^E)\log A})$. Using even more processors, the cost of \textit{expanding} a state can be reduced to the same by generating its successors in parallel. If still more processors are available, some may dedicate to the task of preemptively computing $h^E$ for states that are likely to be generated in the future, e.g. near the A* frontier.

@@

\section{Analysis of HSPr (section to be deleted)}

@@

Another use of E-graphs: if something goes wrong along the path, we can replan using the E-graph to get a faster and less disruptive result.
